---
title: "CosmicWimpout"
output:
  pdf_document: default
  html_document: default
date: "2023-03-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## `roll.dice` function

Rolls a set of dice in Cosmic Wimpout

**Input: **
- n.dice: number of dice to roll
- wild (boolean): whether the wild die is in the set of dice

**Output: **
- dice: set of dice
- wild: the wild die (NULL if wild not rolled)

```{r}
roll.dice <- function(n.dice, wild){
  if(wild){
    n.dice <- n.dice-1
    w <- sample(size=1,c(10,2:6))
    dice <- sample(size=n.dice,c(10,2:6),replace=TRUE)
  }
  else{
    w <- NULL
    dice <- sample(size=n.dice,c(10,2:6),replace=TRUE)
  }
  
  return(list(dice=dice,wild=w))
}
```

```{r}
roll.dice(5,TRUE)
roll.dice(4,FALSE)
```

## `score` function

Scores a roll.

**Input: ** 
- A Comic Wimpout roll (`roll.dice` function defined above)
- Starting score to add on to

**Output: **
- Score: Player score (excludes using wild die as an individual 5/10)
- wild.used: tells player whether the wild die was used to complete a set
- dice.remaining: tells player how many dice they have for their next roll
- must.continue (boolean): tells player if they need to continue to roll.
- triple: if a triple was rolled, which tripple. False otherwise.

```{r}
score <- function(roll,score=0){
  
  wild.used <- FALSE
  w <- roll$wild; if(is.null(w)) wild.used <- TRUE
  r <- roll$dice
  
  dice.remaining <- length(r)
  must.continue <- FALSE
  triple <- FALSE
  
  for(i in c(10,6:2)){
    n <- length(which(r==i))
    
    if(n==4){
      if(w==i|w==3){ # if wild matches i or wild completes 5
        score <- score + i*100
        wild.used <- TRUE
        
        must.continue <- TRUE # all five dice scored
      } else{ # do not use wild
        # score 3/4
        score <- score + i*10
        dice.remaining <- dice.remaining-3
        triple <- i
        must.continue <- TRUE
        
        # if 4th is a 10 or 5
        score <- score + (i==5|i==10)*i
        dice.remaining <- dice.remaining - (i==5|i==10)
      }
    }
    else if(n==3){
      score <- score + i*10
      dice.remaining <- dice.remaining-3
      triple <- i
      must.continue <- TRUE # triple
    }
    else if(n==2){
      if((w==i|w==3)&!wild.used){
        score <- score + i*10
        wild.used <- TRUE
        triple <- i
        must.continue <- TRUE # triple
      }
      else{
        # if 5/10 dice, score them, remove two dice
        score <- score + (i==5|i==10)*i*2 
        dice.remaining <- dice.remaining - (i==5|i==10)*2 
      }
    }
    else if(n==1){
      # if 5/10 die, score it, remove one dice
        score <- score + (i==5|i==10)*i
        dice.remaining <- dice.remaining - (i==5|i==10)
    }
  }
  
  dice.remaining <- dice.remaining + !wild.used
  if(dice.remaining==0) must.continue <- TRUE
  else if(dice.remaining==1){
    if(!wild.used){
      score <- score + w*(w==5|w==10)
      wild.used <- (w==5|w==10)
      dice.remaining <- dice.remaining-(w==5|w==10)
    }
  }
  else if(dice.remaining==5){ # if no dice scored, must score wild
    score <- score + w*(w==5|w==10)
    wild.used <- (w==5|w==10)
    dice.remaining <- dice.remaining-(w==5|w==10)
  }
  
  return(list(score=score,wild.used=wild.used,dice.remaining=dice.remaining,must.continue=must.continue,triple=triple))
}
```

```{r}
roll <- roll.dice(5,wild=TRUE)
roll
```

```{r}
score(roll)
```

## Experiments

```{r}
  roll <- roll.dice(5,TRUE)
  
  s <- score(roll)
  continue <- FALSE
  
  if(s$must.continue){
    continue <- TRUE
  } else{
    continue <- (s$score < 50) # continue if score is less than 50
  }
  
  
  if(s$must.continue){
    roll <- roll.dice(s$dice.remaining,!s$wild.used)
  } else{
    roll <-
  }
```







